---
title: "Main notebook"
output: pdf_document
date: '2023-02-01'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Acknowledgements
The data and a large portion of data preprocessing is provided by Jane Reid. The re-implementation into INLA is also largely based on the work from Stefanie Muff.

# Data loading

```{r data loading}
library(MCMCglmm)
library(MASS)
library(nadiv)
library(bdsmatrix)
library(INLA)
library(QGglmm)
library(SMisc)
library(ggplot2)

qg.data.gg.inds <- read.table("../data/qg.data.gg.inds.steffi.txt", header=T)
d.ped <- ped.prune.inds <- read.table("../data/ped.prune.inds.steffi.txt", header=T)
d.Q <-  read.table("../data/Q.data.steffi.txt", header=T)

qg.data.gg.inds$natalyr.id <- qg.data.gg.inds$natalyr.no

```

Scaling continuous variances can be more stable: (I think?)
```{r}
qg.data.gg.inds$f.coef.sc <- scale(qg.data.gg.inds$f.coef,scale=FALSE)
qg.data.gg.inds$g1.sc <- scale(qg.data.gg.inds$g1,scale=FALSE)
qg.data.gg.inds$natalyr.no.sc <- scale(qg.data.gg.inds$natalyr.no,scale=FALSE)
qg.data.gg.inds$brood.date.sc <- scale(qg.data.gg.inds$brood.date,scale=FALSE)
```

The sex covariate is either `1` or `2`, so we binarize this covariate:
```{r}
qg.data.gg.inds$sex <- qg.data.gg.inds$sex.use.x1 - 1 
```

## Deriving *A*

In order to derive the A matrix, we need to work a bit
```{r}
d.ped <- nadiv::prepPed(d.ped)
```

In particular, for INLA we need ids that run from 1 to the number of individuals
```{r}
d.ped$id <- 1:(nrow(d.ped))
```

Need a map file to keep track of the ids
```{r}
d.map <- d.ped[,c("ninecode","id")]
d.map$g1 <- d.Q[match(d.map$ninecode,d.Q$ninecode),"g1"]
d.map$foc0 <- d.Q[match(d.map$ninecode,d.Q$ninecode),"foc0"]
```

Give mother and father the id
```{r}
d.ped$mother.id <- d.map[match(d.ped$gendam, d.map$ninecode),"id"]
d.ped$father.id <- d.map[match(d.ped$gensire, d.map$ninecode),"id"]
```

Make the inverse A matrix using the `nadiv` package:
```{r}
Cmatrix <- nadiv::makeAinv(d.ped[,c("id","mother.id","father.id")])$Ainv
```

Store the id twice: Once for the breeding value, and once for the independent residuals u with variance 1 (the latter are not going to be included in the end, but we checked what happened when they were there)
```{r}
qg.data.gg.inds$id <- d.map[match(qg.data.gg.inds$ninecode, d.map$ninecode), "id"]
qg.data.gg.inds$u <- 1:nrow(qg.data.gg.inds)
```

# INLA

The general INLA formula is provided below, where `f()` encode the random effect:
```{r}
formula.inla.scaled = surv.ind.to.ad ~ f.coef.sc + g1.sc + natalyr.no.sc + brood.date.sc + sex +
  f(nestrec, model="iid",hyper=list(
    prec=list(initial=log(1/0.05), prior="pc.prec",param=c(1,0.05)) # PC priors
  )) +
  f(natalyr.id, model="iid",hyper=list(
    prec=list(initial=log(1/0.25), prior="pc.prec",param=c(1,0.05)) # PC priors
  )) +
  f(id,model="generic0", # Here we need to specify the covariance matrix 
    Cmatrix=Cmatrix,     #    via the inverse (Cmatrix)
    # constr = TRUE,
    hyper=list(
      prec=list(initial=log(1/10), prior="pc.prec",param=c(1,0.05)) # PC priors
    ))  #+ # this last part is only needed if the u~N(0,1) residuals are included
# f(u, model="iid",
#   constr=TRUE,
#   hyper=list(
#     prec=list(initial=log(1), fixed=TRUE) # Fixed variance to 1
#   )) # This last component adds the independent residuals with variance 1;
```

Now we call INLA models, this is the slowest part:

```{r}

fit.inla.probit = inla(formula=formula.inla.scaled, family="binomial",
                             data=qg.data.gg.inds,
                             control.compute=list(dic=T),
                             control.family = list(link = "probit")
)

fit.inla.logit = inla(formula=formula.inla.scaled, family="binomial",
                             data=qg.data.gg.inds,
                             control.compute=list(dic=T), 
                             control.family = list(link = "logit")
)

fit.inla.gaussian = inla(formula=formula.inla.scaled, family="gaussian",
                             data=qg.data.gg.inds,
                             control.compute=list(dic=T) 
)

```

We define a general method for looking at the results of an INLA model.

```{r}
analyze.inla.fit <- function(inla.fit){
  #' Check if there is a problem (ok of =0) and dic
  inla.fit$mode$mode.status
  inla.fit$dic$dic
  
  summary(inla.fit)
  
  inla.fit$summary.hyperpar
  inla.fit$summary.fixed
  
  inla_mmarginal(inla.fit)
  inla_emarginal(inla.fit)
  
  #' Plotting the posterior marginals for the variances; This is a bit cumbersome,
  #' because INLA works with precisions, so we need a transformation. 
  #' The code below does it for us:
  par(mfrow=c(1,3))
  
  plot(inla.tmarginal(
    function(x) 1/x,inla.fit$marginals.hyperpar$`Precision for nestrec`),
    type="l",main = "next")
  
  plot(inla.tmarginal(
    function(x) 1/x,inla.fit$marginals.hyperpar$`Precision for natalyr.id`),
    type="l",main="natalyr")
  
  plot(inla.tmarginal(
    function(x) 1/x,inla.fit$marginals.hyperpar$`Precision for id`),
    type="l",main="animal")
}

```

Below is a method used to obtain posterior distribution of the back-transformed heritability:
```{r}
get.h2 <- function(inla.fit, n, use.scale=F, model=NA){
  samples <- inla.hyperpar.sample(n=n,inla.fit)

  denominator = 1/samples[,"Precision for id"] + 1/samples[,"Precision for natalyr.id"] +
    1/samples[,"Precision for nestrec"]
  
  if(use.scale){
    # We need model specification to use scale
    stopifnot(model %in% c("binomN.probit", "binomN.logit"))
    scale.param = ifelse(model == "binomN.probit", 1, pi^2/3)
    denominator = denominator + scale.param
  }
  
  h2.inla <- (1/samples[,"Precision for id"]) / denominator
  return(h2.inla)
}

get.h2.from.qgparams <- function(inla.fit, modelname, n, n.obs=nrow(d.ped)){
  #' Computes a posterior of data-scale heritability (h2) using QGParams
  #'  
  #' Params:
  #' inlafit    the fitted INLA object
  #' modelname  a string specifying model
  #'    ("Gaussian", "binomN.probit", "binomN.logit")
  #' n.obs      keyword argument if binomial model is has N != 1 trials
  #'    NB: Should explicitly be set to NULL if not relevant (Gaussian model)
  stopifnot(modelname %in% c("Gaussian", "binomN.probit", "binomN.logit"))
  samples.posterior <- inla.hyperpar.sample(n=n,inla.fit)
  mu = inla.fit$summary.fixed$mean[1]
  va.samples = 1/samples.posterior[,"Precision for id"]
  vp.samples = va.samples +
    1/samples.posterior[,"Precision for natalyr.id"] +
    1/samples.posterior[,"Precision for nestrec"]
  
  kwargs = list(n.obs=n.obs, verbose=F)
  h2.getter = function(...){get("h2.obs", suppressWarnings(QGparams(...)))}
  out = mapply(h2.getter, mu, va.samples, vp.samples, modelname, MoreArgs=kwargs)
  return(out)
}
```

We extend the contents of the INLA fit to include heritabilities on the different scales

```{r}
# Running cell
n.samples = 10000
fit.inla.gaussian$h2.latent = get.h2(fit.inla.gaussian, n.samples)
fit.inla.logit$h2.latent = get.h2(fit.inla.logit, n.samples)
fit.inla.probit$h2.latent = get.h2(fit.inla.probit, n.samples)

fit.inla.logit$h2.scaled = get.h2(fit.inla.logit, n.samples, use.scale=T,
                                  model="binomN.logit")
fit.inla.probit$h2.scaled = get.h2(fit.inla.probit, n.samples,use.scale=T,
                                   model="binomN.probit")

fit.inla.logit$h2.qgglmm  = get.h2.from.qgparams(fit.inla.logit, "binomN.logit", n.samples)
fit.inla.probit$h2.qgglmm = get.h2.from.qgparams(fit.inla.probit, "binomN.probit", n.samples)
```

Plotting some histograms:

Trying one with all overlapping:
```{r}
df.latent.h2 = data.frame(samples=c(unname(fit.inla.logit$h2.latent), unname(fit.inla.probit$h2.latent)),
                      Model=c(
                        rep("Logit",length(fit.inla.logit$h2.latent)),
                        rep("Probit",length(fit.inla.probit$h2.latent))
                      )
                        )

ggplot(df.latent.h2, aes(x=samples, fill=Model)) +
  geom_density(alpha=0.5) +
  ggtitle("Posterior of latent scale heritability for binomial INLA models") +
  xlab("(Latent-scale) Heritability") +
  ylab("Density")
```

```{r}
df.transformed.h2 = data.frame(
  samples=c(
    unname(fit.inla.logit$h2.qgglmm),
    unname(fit.inla.probit$h2.qgglmm),
    unname(fit.inla.gaussian$h2.latent)
    ),
  Model=c(
    rep("Logit",length(fit.inla.logit$h2.qgglmm)),
    rep("Probit", length(fit.inla.probit$h2.qgglmm)),
    rep("Gaussian (no transformation)",length(fit.inla.gaussian$h2.latent))
    )
  )

ggplot(df.transformed.h2, aes(x=samples, fill=Model)) +
  geom_density(alpha=0.5) +
  ggtitle("Posterior of transformed heritability") +
  xlab("(Latent-scale) Heritability") +
  ylab("Density")
```

# Simulation data

```{r}

if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
if (!require("GeneticsPed", quietly = TRUE))
    BiocManager::install("GeneticsPed")

# library("pedigree")
library("GeneticsPed") 
library("MCMCglmm")

# set a value for the Ne/Nc ratio
# here is 0.5 but it can be set to 0.15 or 0.05
NeNc <- 0.5
# set other parameters 
# (idgen = id per generations, nGen = number of generations)
idgen <- 100
nGen <- 9

# generate the pedigree using the function generatePedigree from GeneticsPed
# nFather and nMother are the number of fathers and mothers per generation
# they are generated according to the selected Ne/Nc ratio
ped0 <- generatePedigree(nId = idgen, nGeneration = nGen, 
                         nFather = idgen * NeNc, nMother = idgen * NeNc)  

# N individuals
Nid <- length(ped0$id)


# set correct format for pedigree
pedigree <- ped0[ , c(1,3,2)]
names(pedigree) <- c("id", "dam", "sire")
head(pedigree)
# generate breeding values (always constant)
# using the function rbv from MCMCglmm
# fix the additive variance value, here is 0.3 but it can be any value
Z <- diag(Nid)
sigmaA <- 0.4
# pedigree <- ped0[ ,1:3]
u <- rbv(pedigree, sigmaA) # Breeding values for each individual 
## We get: the condition has length > 1 and only the first element will be used #??? 

# How do we generate the rest of the data tho? Like number survived and whatnot
```

Now we try to run it through the model pipeline:
```{r}
simulated.d.ped <- pedigree
simulated.d.ped <- nadiv::prepPed(simulated.d.ped)
simulated.Cmatrix <- nadiv::makeAinv(simulated.d.ped)$Ainv # TODO might need to index here?

``` 