\chapter*{Acknowledgements}


First and foremost, I wish to start by expressing my gratitude to my supervisor, Associate Professor Stefanie Muff, for the exceptional guidance and support I received throughout the entire master's thesis process. From the very start, she has helped me get started with both the mathematical background and especially the biological background, the latter of which I had almost no prior knowledge. The feedback and discussions that we had during the semester, the frequent meetings and the fast response times have truly been invaluable in shaping my thesis.

Additionally, I would like to thank those who provided me with important methodology and data making this master's thesis possible. Thanks to Professor Jane Reid for allowing me to work with her dataset for my research, as it has been instrumental both in producing the results and in helping me understand the real-world applications. I would also like to thank Pierre de Villemereuil for publishing important methodologies that laid the foundations of this thesis.

Lastly, as this master's thesis concludes my five-year journey at NTNU and my M.Sc. in Industrial Mathematics, I would like to thank some people who have helped me over the past years. Thank you to wonderful friends and family for continued support and motivation. I especially thank my fiancé Lars Dahlgren for standing by my side through thick and thin and for much-needed encouragement over the last five years. Thank you to all the student organizations I have had the opportunity to be a part of, especially the Nabla student organization and the International Student Festival in Trondheim (ISFiT). The skills and competence I have gained from these organizations and the students I have met there have been formative in shaping who I am and taught me how to tackle prior and upcoming challenges of all calibers.